---
title: "My R Markdown"
author: "Virginia Liebrecht"
date: "2025-11-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

R Markdown is a powerful format that combines text, R code and output (such as tables or plots) in a single, reproducible document.
It uses a simple Markdown syntax for formatting text and integrates R code chunks to automatically include data analysis outputs. 

When you click the Knit button in RStudio, the document is rendered into HTML, PDF, or Word, and blends code, text, and visuals seamlessly.


## Example 1: Data Summary
Here is an example, using the built in cars dataset, which records the speed of cars and their stopping distance. 

```{r cars}
summary(cars)
```

The summary shows descriptive statistics such as the minimum, maximum, mean and quartiles for each variable. 

## Markdown and LaTex Examples 

You can include formatted text and mathematical expressions easily in R Markdown
Inline math example: $E = mc^2$

Displayed equation example (Quadratic Formula):
$$
x = \frac{-b \pm \sqrt{b^2 - 4ac}}{2a}
$$
This shows how LaTeX can render fractions, square roots, and symbols beautifully in reports.    
    
## Example 2: Including Plots
Here is an example of a basic plot using the built-in pressure dataset:

```{r pressure, echo=FALSE}
plot(pressure,
     main = "Pressure vs Temperature",
     xlab = "Temperature (C)",
     ylab = "Pressure (mm of Hg)",
     col = "blue",
     pch = 19)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


## Reflection
Using R Markdown makes it simple to combine narrative, code, and results in one document. 
It's especially useful for reproducible research - if your data or code changes, the whole document updates automatically.

What I learned:
- How Markdown syntax and LaTeX math work together.
- How code chunks display both input and output.
- The benefit of knitting to HTML for visual clarity.

Overall, R Markdown is a great tool for blending text and computation, which is much easier than copying results manually into reports. 
